---
title: 'Ex3: Data Wrangling'
author: "mannava praneeth chowdary"
date: "02/05/2025"
output:
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exploring data sets: using dplyr on external data

-   An important part about being a data scientist is asking questions. Write a question you may be interested in about the data set `data/nba_teams_2016.csv`, and then use dplyr to figure out the answer! 
if i want to find a specific number in huge dataset even if it exists in multiple places, how do i do it?
example based on nba dataset
# # Define the number you are searching for
search_value <- 9421

# Find all occurrences of the number in any column
matches <- nba_teams == search_value

# Display the rows where the value is found
nba_teams[rowSums(matches, na.rm = TRUE) > 0, ]


-   Load the `dplyr` library

```{r}
library(dplyr)
```

-   Use the `read.csv()` function to read in the included data set. Remember to save it as a variable.

```{r}
nba_teams <- read.csv("/Users/praneeth/Desktop/r cls/data/nba_teams_2016.csv")
```

-   View the data frame you loaded, and get some basic information about the number of rows/columns.Note the "X" preceding some of the column titles as well as the "\*" following the names of teams that made it to the playoffs that year.

```{r}
str(nba_teams)
num_rows <- nrow(nba_teams)
num_cols <- ncol(nba_teams)
cat("The dataset has", num_rows, "rows and", num_cols, "columns.\n")
colnames(nba_teams)
```

-   Add a column that gives the turnovers to steals ratio (TOV / STL) for each team.

```{r}
nba_teams <- nba_teams %>%
  mutate(TOV_STL_Ratio = TOV / STL)
```

-   Sort the teams from lowest turnover/steal ratio to highest. Which team has the lowest turnover/steal ratio?

```{r}
nba_sorted <- nba_teams %>%
  arrange(TOV_STL_Ratio)
lowest_tov_stl_team <- nba_sorted[1, "Team"]
print(lowest_tov_stl_team)
```

-   Using the pipe operator, create a new column of assists per game (AST / G), AND sort the data.frame by this new column in descending order.

```{r}
nba_teams <- nba_teams %>%
  mutate(AST_per_Game = AST / G) %>%
  arrange(desc(AST_per_Game))
```

-   Create a data frame called `good_offense` of teams that scored more than 8700 points (PTS) in the season

```{r}
good_offense <- nba_teams %>%
  filter(PTS > 8700)
```

-   Create a data frame called `good_defense` of teams that had more than 470 blocks (BLK)

```{r}
ood_defense <- nba_teams %>%
  filter(BLK > 470)
```

-   Create a data frame called `offense_stats` that only shows offensive rebounds (ORB), field-goal % (FG.), and assists (AST) along with the team name.

```{r}
offense_stats <- nba_teams %>%
  select(Team, ORB, FG., AST)
```

-   Create a data frame called `defense_stats` that only shows defensive rebounds (DRB), steals (STL), and blocks (BLK) along with the team name.

```{r}
defense_stats <- nba_teams %>%
  select(Team, DRB, STL, BLK)
```

-   Create a function called `better_shooters` that takes in two teams and returns a data frame of the team with the better field-goal percentage. Include the team name, field-goal percentage, and total points in your resulting data frame.

```{r}
better_shooters <- function(team1, team2) {
  nba_teams %>%
    filter(Team %in% c(team1, team2)) %>%
    select(Team, FG., PTS) %>%
    arrange(desc(FG.))
}
```

-   Call the function on two teams to compare them (remember the `*` if needed)

```{r}
better_shooters("Golden State Warriors*", "San Antonio Spurs*")
```

```{r}
head(nba_teams)  # View first few rows
tail(nba_teams)  # View last few rows
print(nba_teams) # Print full dataset (not recommended for large datasets)
summary(nba_teams)  # Get summary statistics

```



## Analyzing data with the `tidyr` package

-   Load necessary packages (`tidyr`, `dplyr`, and `ggplot2`)

```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
```

-   Load the `data/avocado.csv` file into a variable `avocados`. Make sure strings are *not* read in as factors

```{r}
avocados <- read.csv("/Users/praneeth/Desktop/r cls/data/avocado.csv", stringsAsFactors = FALSE)
```

-   To tell R to treat the `Date` column as a date (not just a string). Redefine that column as a date using the `as.Date()` function (hint: use the `mutate` function)

```{r}
avocados <- avocados %>%
  mutate(Date = as.Date(Date))
```

-   The file had some uninformative column names, so rename these columns: -- `X4046` to `small_haas` -- `X4225` to `large_haas` -- `X4770` to `xlarge_haas`

```{r}
avocados <- avocados %>%
  rename(
    small_haas = X4046,
    large_haas = X4225,
    xlarge_haas = X4770
  )
```

-   The data only has sales for haas avocados. Create a new column `other_avos` that is the Total. Volume minus all haas avocados (small, large, xlarge)

```{r}
avocados <- avocados %>%
  mutate(other_avos = Total.Volume - small_haas - large_haas - xlarge_haas)
```

-   To perform analysis by avocado size, create a dataframe `by_size` that has only `Date`, `other_avos`, `small_haas`, `large_haas`, `xlarge_haas`

```{r}
by_size <- avocados %>%
  select(Date, other_avos, small_haas, large_haas, xlarge_haas)
```

-   In order to visualize this data, it needs to be reshaped. The four columns `other_avos`, `small_haas`, `large_haas`, `xlarge_haas` need to be **gathered** together into a single column called `size`. The volume of sales (currently stored in each column) should be stored in a new column called `volume`. Create a new dataframe `size_gathered` by passing the `by_size` data frame to the `gather()` function. `size_gathered` will only have 3 columns: `Date`, `size`, and `volume`.

```{r}
size_gathered <- by_size %>%
  gather(key = "size", value = "volume", other_avos, small_haas, large_haas, xlarge_haas)
```

-   Using `size_gathered`, compute the average sales volume of each size (hint, first `group_by` size, then compute using `summarize`)

```{r}
avg_sales_by_size <- size_gathered %>%
  group_by(size) %>%
  summarize(avg_volume = mean(volume, na.rm = TRUE))
```

-   Investigate sales by avocado type (conventional, organic). Create a new data frame `by_type` by grouping the `avocados` dataframe by `Date` and `type`, and calculating the sum of the `Total.Volume` for that type in that week (resulting in a data frame with 2 rows per week).

```{r}
by_type <- avocados %>%
  group_by(Date, type) %>%
  summarize(total_volume = sum(Total.Volume, na.rm = TRUE))
```

```{r}
colnames(avocados)
summary(avocados)
str(avocados)
```

